Embracing the Laws of Physics: \\ Three Reversible Models of Computation

### 2018-01-01 survey/survey.tex
consistently use odot for first level composition. Change naming in first example in Data III to avoid matching A with 2 and B with 1, which is just confusing (even if correct).

\begin{figure}[t]
Let $c_1 : t_1 \leftrightarrow t_2$,  $c_2 : t_2 \leftrightarrow t₃$, and $c_3 : t_3 \leftrightarrow t_4$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {c_1 \circledcirc (c_2 \circledcirc c_3) \Leftrightarrow (c_1 \circledcirc c_2) \circledcirc c_3}
\\
  {(c_1 \oplus (c_2 \oplus c_3)) \circledcirc \assoclp \Leftrightarrow \assoclp \circledcirc ((c_1 \oplus c_2) \oplus c_3)}
\\
  {(c_1 \otimes (c_2 \otimes c_3)) \circledcirc \assoclt \Leftrightarrow \assoclt \circledcirc ((c_1 \otimes c_2) \otimes c_3)}
\\
  {((c_1 \oplus c_2) \oplus c_3) \circledcirc \assocrp \Leftrightarrow \assocrp \circledcirc (c_1 \oplus (c_2 \oplus c_3))}
\\
  {((c_1 \otimes c_2) \otimes c_3) \circledcirc \assocrt \Leftrightarrow \assocrt \circledcirc (c_1 \otimes (c_2 \otimes c_3))}
\\
  {\assocrp \circledcirc \assocrp \Leftrightarrow ((\assocrp \oplus \idc) \circledcirc \assocrp) \circledcirc (\idc \oplus \assocrp)}
\\
  {\assocrt \circledcirc \assocrt \Leftrightarrow ((\assocrt \otimes \idc) \circledcirc \assocrt) \circledcirc (\idc \otimes \assocrt)}
\end{array}\]
\caption{\label{figj}Signatures of level-1 $\Pi$-combinators: associativity}
\end{figure}

\begin{figure}[t]
Let $a : t_1 \leftrightarrow t_2$, $b : t_3 \leftrightarrow t_4$, and $c : t₅ \leftrightarrow t₆$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {((a \oplus b) \otimes c) \circledcirc \dist \Leftrightarrow \dist \circledcirc ((a \otimes c) \oplus (b \otimes c))}
\\
  {(a \otimes (b \oplus c)) \circledcirc \distl \Leftrightarrow \distl \circledcirc ((a \otimes b) \oplus (a \otimes c))}
\\
  {((a \otimes c) \oplus (b \otimes c)) \circledcirc \factor \Leftrightarrow \factor \circledcirc ((a \oplus b) \otimes c)}
\\
  {((a \otimes b) \oplus (a \otimes c)) \circledcirc \factorl \Leftrightarrow \factorl \circledcirc (a \otimes (b \oplus c))}
\end{array}\]
\caption{\label{figi}Signatures of level-1 $\Pi$-combinators: distributivity and factoring}
\end{figure}

\begin{figure}[t]
Let $c, c_1, c_2, c_3 : t_1 \leftrightarrow t_2$ and $c', c'' : t_3 \leftrightarrow t_4$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\idc \circledcirc \, c \Leftrightarrow c}
\quad
  {c \, \circledcirc \idc \, \Leftrightarrow c}
\quad
  {c\,\, \circledcirc\,! c \Leftrightarrow \idc}
\quad
  {! c \circledcirc c \Leftrightarrow \idc}
\\
  {c \Leftrightarrow c}
\quad
\Rule{}
  {c_1 \Leftrightarrow c_2 \quad c_2 \Leftrightarrow c_3}
  {c_1 \Leftrightarrow c_3}
  {}
\quad
\Rule{}
  {c_1 \Leftrightarrow c' \quad c_2 \Leftrightarrow c''}
  {c_1 \circledcirc c_2 \Leftrightarrow c' \circledcirc c''}
  {}
\end{array}\]
\caption{\label{figh}Signatures of level-1 $\Pi$-combinators: identity and composition}
\end{figure}

\begin{figure}[t]
Let $c₀ : 0 \leftrightarrow 0$, $c_1 : 1 \leftrightarrow 1$, and $c : t_1 \leftrightarrow t_2$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\identlp \circledcirc c \Leftrightarrow (c₀ \oplus c) \circledcirc \identlp}
\qquad
  {\identrp \circledcirc (c₀ \oplus c) \Leftrightarrow c \circledcirc \identrp}
\\
  {\identlsp \circledcirc c \Leftrightarrow (c \oplus c₀) \circledcirc \identlsp}
\qquad
  {\identrsp \circledcirc (c \oplus c₀) \Leftrightarrow c \circledcirc \identrsp}
\\
  {\identlt \circledcirc c \Leftrightarrow (c_1 \otimes c) \circledcirc \identlt}
\qquad
  {\identrt \circledcirc (c_1 \otimes c) \Leftrightarrow c \circledcirc \identrp}
\\
  {\identlst \circledcirc c \Leftrightarrow (c \otimes c_1) \circledcirc \identlst}
\qquad
  {\identrst \circledcirc (c \otimes c_1) \Leftrightarrow c \circledcirc \identrst}
\\
  {\identlt \Leftrightarrow \distl \circledcirc (\identlt \oplus \identlt)}
\\
\identlp \Leftrightarrow \swapp \circledcirc \identrsp
\qquad
\identlt \Leftrightarrow \swapt \circledcirc \identrst
\end{array}\]
\caption{\label{figg}Signatures of level-1 $\Pi$-combinators: unit}
\end{figure}

\begin{figure}[t]
Let $c_1 : t_1 \leftrightarrow t_2$ and $c_2 : t_3 \leftrightarrow t_4$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\swapp \circledcirc (c_1 \oplus c_2) \Leftrightarrow (c_2 \oplus c_1) \circledcirc \swapp}
\quad
  {\swapt \circledcirc (c_1 \otimes c_2) \Leftrightarrow (c_2 \otimes c_1) \circledcirc \swapt}
\\
  {(\assocrp \circledcirc \swapp) \circledcirc \assocrp \Leftrightarrow ((\swapp \oplus \idc) \circledcirc \assocrp) \circledcirc (\idc \oplus \swapp)}
\\
  {(\assoclp \circledcirc \swapp) \circledcirc \assoclp \Leftrightarrow ((\idc \oplus \swapp) \circledcirc \assoclp) \circledcirc (\swapp \oplus \idc)}
\\
  {(\assocrt \circledcirc \swapt) \circledcirc \assocrt \Leftrightarrow ((\swapt \otimes \idc) \circledcirc \assocrt) \circledcirc (\idc \otimes \swapt)}
\\
  {(\assoclt \circledcirc \swapt) \circledcirc \assoclt \Leftrightarrow ((\idc \otimes \swapt) \circledcirc \assoclt) \circledcirc (\swapt \otimes \idc)}
\end{array}\]
\caption{\label{figf}Signatures of level-1 $\Pi$-combinators: commutativity and associativity}
\end{figure}

\begin{figure}[t]
Let $c_1 : t_1 \leftrightarrow t_2$, $c_2 : t_3 \leftrightarrow t_4$, $c_3 : t_1 \leftrightarrow t_2$, and $c_4 : t_3 \leftrightarrow t_4$. \\
Let $a_1 : t₅ \leftrightarrow t_1$,  $a_2 : t₆ \leftrightarrow t_2$, $a_3 : t_1 \leftrightarrow t_3$, and $a_4 : t_2 \leftrightarrow t_4$.
\[\def\arraystretch{1.3}
\begin{array}{c}
\Rule{}
  {c_1 \Leftrightarrow c_3 \quad c_2 \Leftrightarrow c_4}
  {c_1 \oplus c_2 \Leftrightarrow c_3 \oplus c_4}
  {}
\qquad
\Rule{}
  {c_1 \Leftrightarrow c_3 \quad c_2 \Leftrightarrow c_4}
  {c_1 \otimes c_2 \Leftrightarrow c_3 \otimes c_4}
  {}
\\
  {\idc \oplus \, \idc \, \Leftrightarrow \idc}
\qquad
  {\idc \otimes \, \idc \, \Leftrightarrow \idc}
\\
  {(a_1 \circledcirc a_3) \oplus (a_2 \circledcirc a_4) \Leftrightarrow (a_1 \oplus a_2) \circledcirc (a_3 \oplus a_4)}
\\
  {(a_1 \circledcirc a_3) \otimes (a_2 \circledcirc a_4) \Leftrightarrow (a_1 \otimes a_2) \circledcirc (a_3 \otimes a_4)}
\end{array}\]
\caption{\label{fige}Signatures of level-1 $\Pi$-combinators: functors}
\end{figure}

\begin{figure}[t]
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\identlsp \oplus \idc ~\Leftrightarrow~ \assocrp \circledcirc (\idc \oplus \, \identlp)}
\\
  {\identlst \otimes \idc ~\Leftrightarrow~ \assocrt \circledcirc (\idc \otimes \, \identlt)}
\end{array}\]
\caption{\label{figd}Signatures of level-1 $\Pi$-combinators: unit and associativity}
\end{figure}

\begin{figure}[t]
Let $c : t_1 \leftrightarrow t_2$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {(c \otimes \idc) \circledcirc \absorbl \Leftrightarrow \absorbl \circledcirc \idc}
\quad
  {(\idc \, \otimes c) \circledcirc \absorbr \Leftrightarrow \absorbr \circledcirc \idc}
\\
  {\idc \circledcirc \, \factorzl \Leftrightarrow \factorzl \circledcirc (\idc \otimes c)}
\quad
  {\idc \circledcirc \, \factorzr \Leftrightarrow \factorzr \circledcirc (c \otimes \idc)}
\\
  {\absorbr \Leftrightarrow \absorbl}
\\
  {\absorbr \Leftrightarrow (\distl \circledcirc (\absorbr \oplus \absorbr)) \circledcirc \identlp}
\\
  {\identlst \Leftrightarrow \absorbr}
\qquad
  {\absorbl \Leftrightarrow \swapt \circledcirc \absorbr}
\\
  {\absorbr \Leftrightarrow (\assoclt \circledcirc (\absorbr \otimes \idc)) \circledcirc \absorbr}
\\
  {(\idc \otimes \absorbr) \circledcirc \absorbl \Leftrightarrow (\assoclt \circledcirc (\absorbl \otimes \idc)) \circledcirc \absorbr}
\\
  {\idc \otimes \, \identlp \Leftrightarrow (\distl \circledcirc (\absorbl \oplus \idc)) \circledcirc \identlp}
\end{array}\]
\caption{\label{figc}Signatures of level-1 $\Pi$-combinators: zero}
\end{figure}

\begin{figure}[t]
\[\def\arraystretch{1.3}
\begin{array}{c}
  {((\assoclp \otimes \idc) \circledcirc \dist) \circledcirc (\dist \oplus \idc) \Leftrightarrow (\dist \circledcirc (\idc \oplus \dist)) \circledcirc \assoclp}
\\
  {\assoclt \circledcirc \distl \Leftrightarrow ((\idc \otimes \distl) \circledcirc \distl) \circledcirc (\assoclt \oplus \assoclt)}
\end{array}\]
\vspace{ -0.5em}
\[\def\arraystretch{1.3}
\begin{array}{rcl}
  (\distl \circledcirc (\dist \oplus \dist)) \circledcirc \assoclp &\Leftrightarrow&
   \dist \circledcirc (\distl \oplus \distl) \circledcirc \assoclp ~\circledcirc \\
&& (\assocrp \oplus \idc) ~\circledcirc \\
&& ((\idc \oplus \swapp) \oplus \idc) ~\circledcirc \\
&&      (\assoclp \oplus \idc)
\end{array}\]
\caption{\label{figb}Signatures of level-1 $\Pi$-combinators: associativity and distributivity}
\end{figure}

\begin{figure}[t]
\[\def\arraystretch{1.3}
\begin{array}{rcl}
  (\idc \otimes \swapp) \circledcirc \distl &\Leftrightarrow& \distl \circledcirc \swapp
\\
  \dist \circledcirc (\swapt \oplus \swapt) &\Leftrightarrow & \swapt \circledcirc \distl
\end{array}\]
\caption{\label{figa}Signatures of level-1 $\Pi$-combinators: commutativity and distributivity}
\end{figure}

@@

In the previous sections, we examined equivalences between
conventional data structures, i.e, sets of values and structured trees
of values. We now consider a richer, less conventional, but
foundational notion of data: programs themselves. Indeed, universal
computation models fundamentally rely on the fact that \emph{programs
  are (or can be encoded as) data}, e.g., a Turing machine can be
encoded as a string that another Turing machine (or even the same
machine) can manipulate. In our setting, the programs developed in the
previous section are reversible deformations between structured finite
types. We now ask whether these reversible deformations can themselves
be subject to (higher-level) reversible deformations?

@@

\begin{center}
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.8}]
  \draw[>=latex,<->,double,red,thick] (2.25,-1.2) -- (2.25,-2.9) ;
%%  \node at (3.3,-2) {$\mathit{swapl}_{+\Leftrightarrow}$} ;
%%  \node at (2.5,-1.3) {$((c_2~\oplus~c_1)~\circledcirc~\mathit{swap}_{+})$};
%%  \node at (2.5,-2.7) {$(\mathit{swap}_{+}~\circledcirc~(c_1~\oplus~c_2))$};
  \draw (-2,-2) ellipse (0.5cm and 1cm);
  \draw[fill] (-2,-1.5) circle [radius=0.025];
  \node[below] at (-2.1,-1.5) {$A$};
  \draw[fill] (-2,-2.5) circle [radius=0.025];
  \node[below] at (-2.1,-2.5) {$B$};

@@

  \node at (3,-3.5) {$c_1$};
  \node at (3,-4.5) {$c_2$};

  \node at (1.5,0.5) {$c_2$};
  \node at (1.5,-0.5) {$c_1$};

@@

\end{tikzpicture}
\end{center}
The top path is the reversible deformation
$(c_2~\oplus~c_1)~\circledcirc~\mathit{swap}_{+}$ which deforms the
type $A$ by $c_2$, deforms the type $B$ by $c_1$, and deforms the
resulting space by a twist that exchanges the two injections into the
sum type. The bottom path performs the twist first and then deforms
the type $A$ by $c_2$ and the type $B$ by $c_1$ as before. If one
could imagine the paths are physical wires and the deformations $c_1$
and $c_2$ as arbitrary deformations on these wires then, holding the
points $A$, $B$, $C$, and $D$ fixed, it is possible to rotate the top
part of the diagram to become identical to the bottom one. In other
words, there exists a higher-level deformation of the program
$(c_2~\oplus~c_1)~\circledcirc~\mathit{swap}_{+}$ to the program
$\mathit{swap}_{+} \circledcirc (c_1~\oplus~c_2)$.

>>>>

\begin{figure}[t]
Let $c_1 : t_1 \leftrightarrow t_2$,  $c_2 : t_2 \leftrightarrow t₃$, and $c_3 : t_3 \leftrightarrow t_4$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {c_1 \odot (c_2 \odot c_3) \Leftrightarrow (c_1 \odot c_2) \odot c_3}
\\
  {(c_1 \oplus (c_2 \oplus c_3)) \odot \assoclp \Leftrightarrow \assoclp \odot ((c_1 \oplus c_2) \oplus c_3)}
\\
  {(c_1 \otimes (c_2 \otimes c_3)) \odot \assoclt \Leftrightarrow \assoclt \odot ((c_1 \otimes c_2) \otimes c_3)}
\\
  {((c_1 \oplus c_2) \oplus c_3) \odot \assocrp \Leftrightarrow \assocrp \odot (c_1 \oplus (c_2 \oplus c_3))}
\\
  {((c_1 \otimes c_2) \otimes c_3) \odot \assocrt \Leftrightarrow \assocrt \odot (c_1 \otimes (c_2 \otimes c_3))}
\\
  {\assocrp \odot \assocrp \Leftrightarrow ((\assocrp \oplus \idc) \odot \assocrp) \odot (\idc \oplus \assocrp)}
\\
  {\assocrt \odot \assocrt \Leftrightarrow ((\assocrt \otimes \idc) \odot \assocrt) \odot (\idc \otimes \assocrt)}
\end{array}\]
\caption{\label{figj}Signatures of level-1 $\Pi$-combinators: associativity}
\end{figure}

\begin{figure}[t]
Let $a : t_1 \leftrightarrow t_2$, $b : t_3 \leftrightarrow t_4$, and $c : t₅ \leftrightarrow t₆$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {((a \oplus b) \otimes c) \odot \dist \Leftrightarrow \dist \odot ((a \otimes c) \oplus (b \otimes c))}
\\
  {(a \otimes (b \oplus c)) \odot \distl \Leftrightarrow \distl \odot ((a \otimes b) \oplus (a \otimes c))}
\\
  {((a \otimes c) \oplus (b \otimes c)) \odot \factor \Leftrightarrow \factor \odot ((a \oplus b) \otimes c)}
\\
  {((a \otimes b) \oplus (a \otimes c)) \odot \factorl \Leftrightarrow \factorl \odot (a \otimes (b \oplus c))}
\end{array}\]
\caption{\label{figi}Signatures of level-1 $\Pi$-combinators: distributivity and factoring}
\end{figure}

\begin{figure}[t]
Let $c, c_1, c_2, c_3 : t_1 \leftrightarrow t_2$ and $c', c'' : t_3 \leftrightarrow t_4$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\idc \odot \, c \Leftrightarrow c}
\quad
  {c \, \odot \idc \, \Leftrightarrow c}
\quad
  {c\,\, \odot\,! c \Leftrightarrow \idc}
\quad
  {! c \odot c \Leftrightarrow \idc}
\\
  {c \Leftrightarrow c}
\quad
\Rule{}
  {c_1 \Leftrightarrow c_2 \quad c_2 \Leftrightarrow c_3}
  {c_1 \Leftrightarrow c_3}
  {}
\quad
\Rule{}
  {c_1 \Leftrightarrow c' \quad c_2 \Leftrightarrow c''}
  {c_1 \odot c_2 \Leftrightarrow c' \odot c''}
  {}
\end{array}\]
\caption{\label{figh}Signatures of level-1 $\Pi$-combinators: identity and composition}
\end{figure}

\begin{figure}[t]
Let $c₀ : 0 \leftrightarrow 0$, $c_1 : 1 \leftrightarrow 1$, and $c : t_1 \leftrightarrow t_2$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\identlp \odot c \Leftrightarrow (c₀ \oplus c) \odot \identlp}
\qquad
  {\identrp \odot (c₀ \oplus c) \Leftrightarrow c \odot \identrp}
\\
  {\identlsp \odot c \Leftrightarrow (c \oplus c₀) \odot \identlsp}
\qquad
  {\identrsp \odot (c \oplus c₀) \Leftrightarrow c \odot \identrsp}
\\
  {\identlt \odot c \Leftrightarrow (c_1 \otimes c) \odot \identlt}
\qquad
  {\identrt \odot (c_1 \otimes c) \Leftrightarrow c \odot \identrp}
\\
  {\identlst \odot c \Leftrightarrow (c \otimes c_1) \odot \identlst}
\qquad
  {\identrst \odot (c \otimes c_1) \Leftrightarrow c \odot \identrst}
\\
  {\identlt \Leftrightarrow \distl \odot (\identlt \oplus \identlt)}
\\
\identlp \Leftrightarrow \swapp \odot \identrsp
\qquad
\identlt \Leftrightarrow \swapt \odot \identrst
\end{array}\]
\caption{\label{figg}Signatures of level-1 $\Pi$-combinators: unit}
\end{figure}

\begin{figure}[t]
Let $c_1 : t_1 \leftrightarrow t_2$ and $c_2 : t_3 \leftrightarrow t_4$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\swapp \odot (c_1 \oplus c_2) \Leftrightarrow (c_2 \oplus c_1) \odot \swapp}
\quad
  {\swapt \odot (c_1 \otimes c_2) \Leftrightarrow (c_2 \otimes c_1) \odot \swapt}
\\
  {(\assocrp \odot \swapp) \odot \assocrp \Leftrightarrow ((\swapp \oplus \idc) \odot \assocrp) \odot (\idc \oplus \swapp)}
\\
  {(\assoclp \odot \swapp) \odot \assoclp \Leftrightarrow ((\idc \oplus \swapp) \odot \assoclp) \odot (\swapp \oplus \idc)}
\\
  {(\assocrt \odot \swapt) \odot \assocrt \Leftrightarrow ((\swapt \otimes \idc) \odot \assocrt) \odot (\idc \otimes \swapt)}
\\
  {(\assoclt \odot \swapt) \odot \assoclt \Leftrightarrow ((\idc \otimes \swapt) \odot \assoclt) \odot (\swapt \otimes \idc)}
\end{array}\]
\caption{\label{figf}Signatures of level-1 $\Pi$-combinators: commutativity and associativity}
\end{figure}

\begin{figure}[t]
Let $c_1 : t_1 \leftrightarrow t_2$, $c_2 : t_3 \leftrightarrow t_4$, $c_3 : t_1 \leftrightarrow t_2$, and $c_4 : t_3 \leftrightarrow t_4$. \\
Let $a_1 : t₅ \leftrightarrow t_1$,  $a_2 : t₆ \leftrightarrow t_2$, $a_3 : t_1 \leftrightarrow t_3$, and $a_4 : t_2 \leftrightarrow t_4$.
\[\def\arraystretch{1.3}
\begin{array}{c}
\Rule{}
  {c_1 \Leftrightarrow c_3 \quad c_2 \Leftrightarrow c_4}
  {c_1 \oplus c_2 \Leftrightarrow c_3 \oplus c_4}
  {}
\qquad
\Rule{}
  {c_1 \Leftrightarrow c_3 \quad c_2 \Leftrightarrow c_4}
  {c_1 \otimes c_2 \Leftrightarrow c_3 \otimes c_4}
  {}
\\
  {\idc \oplus \, \idc \, \Leftrightarrow \idc}
\qquad
  {\idc \otimes \, \idc \, \Leftrightarrow \idc}
\\
  {(a_1 \odot a_3) \oplus (a_2 \odot a_4) \Leftrightarrow (a_1 \oplus a_2) \odot (a_3 \oplus a_4)}
\\
  {(a_1 \odot a_3) \otimes (a_2 \odot a_4) \Leftrightarrow (a_1 \otimes a_2) \odot (a_3 \otimes a_4)}
\end{array}\]
\caption{\label{fige}Signatures of level-1 $\Pi$-combinators: functors}
\end{figure}

\begin{figure}[t]
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\identlsp \oplus \idc ~\Leftrightarrow~ \assocrp \odot (\idc \oplus \, \identlp)}
\\
  {\identlst \otimes \idc ~\Leftrightarrow~ \assocrt \odot (\idc \otimes \, \identlt)}
\end{array}\]
\caption{\label{figd}Signatures of level-1 $\Pi$-combinators: unit and associativity}
\end{figure}

\begin{figure}[t]
Let $c : t_1 \leftrightarrow t_2$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {(c \otimes \idc) \odot \absorbl \Leftrightarrow \absorbl \odot \idc}
\quad
  {(\idc \, \otimes c) \odot \absorbr \Leftrightarrow \absorbr \odot \idc}
\\
  {\idc \odot \, \factorzl \Leftrightarrow \factorzl \odot (\idc \otimes c)}
\quad
  {\idc \odot \, \factorzr \Leftrightarrow \factorzr \odot (c \otimes \idc)}
\\
  {\absorbr \Leftrightarrow \absorbl}
\\
  {\absorbr \Leftrightarrow (\distl \odot (\absorbr \oplus \absorbr)) \odot \identlp}
\\
  {\identlst \Leftrightarrow \absorbr}
\qquad
  {\absorbl \Leftrightarrow \swapt \odot \absorbr}
\\
  {\absorbr \Leftrightarrow (\assoclt \odot (\absorbr \otimes \idc)) \odot \absorbr}
\\
  {(\idc \otimes \absorbr) \odot \absorbl \Leftrightarrow (\assoclt \odot (\absorbl \otimes \idc)) \odot \absorbr}
\\
  {\idc \otimes \, \identlp \Leftrightarrow (\distl \odot (\absorbl \oplus \idc)) \odot \identlp}
\end{array}\]
\caption{\label{figc}Signatures of level-1 $\Pi$-combinators: zero}
\end{figure}

\begin{figure}[t]
\[\def\arraystretch{1.3}
\begin{array}{c}
  {((\assoclp \otimes \idc) \odot \dist) \odot (\dist \oplus \idc) \Leftrightarrow (\dist \odot (\idc \oplus \dist)) \odot \assoclp}
\\
  {\assoclt \odot \distl \Leftrightarrow ((\idc \otimes \distl) \odot \distl) \odot (\assoclt \oplus \assoclt)}
\end{array}\]
\vspace{ -0.5em}
\[\def\arraystretch{1.3}
\begin{array}{rcl}
  (\distl \odot (\dist \oplus \dist)) \odot \assoclp &\Leftrightarrow&
   \dist \odot (\distl \oplus \distl) \odot \assoclp ~\odot \\
&& (\assocrp \oplus \idc) ~\odot \\
&& ((\idc \oplus \swapp) \oplus \idc) ~\odot \\
&&      (\assoclp \oplus \idc)
\end{array}\]
\caption{\label{figb}Signatures of level-1 $\Pi$-combinators: associativity and distributivity}
\end{figure}

\begin{figure}[t]
\[\def\arraystretch{1.3}
\begin{array}{rcl}
  (\idc \otimes \swapp) \odot \distl &\Leftrightarrow& \distl \odot \swapp
\\
  \dist \odot (\swapt \oplus \swapt) &\Leftrightarrow & \swapt \odot \distl
\end{array}\]
\caption{\label{figa}Signatures of level-1 $\Pi$-combinators: commutativity and distributivity}
\end{figure}

@@

In the previous sections, we examined equivalences between
conventional data structures, i.e, sets of values and structured trees
of values. We now consider a richer, less conventional, but
foundational notion of data: programs themselves. Indeed, universal
computation models fundamentally rely on the fact that \emph{programs
are (or can be encoded as) data}, e.g., a Turing machine can be
encoded as a string that another Turing machine (or even the same
machine) can manipulate. In our setting, the programs developed in the
previous section are reversible deformations between structured finite
types. We now ask whether these programs can themselves
be subject to (higher-level) reversible deformations?

@@

\begin{center}
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.8}]
  \draw[>=latex,<->,double,red,thick] (2.25,-1.2) -- (2.25,-2.9) ;
%%  \node at (3.3,-2) {$\mathit{swapl}_{+\Leftrightarrow}$} ;
%%  \node at (2.5,-1.3) {$((c_2~\oplus~c_1)~\odot~\mathit{swap}_{+})$};
%%  \node at (2.5,-2.7) {$(\mathit{swap}_{+}~\odot~(c_1~\oplus~c_2))$};
  \draw (-2,-2) ellipse (0.5cm and 1cm);
  \draw[fill] (-2,-1.5) circle [radius=0.025];
  \node[below] at (-2.1,-1.5) {$A$};
  \draw[fill] (-2,-2.5) circle [radius=0.025];
  \node[below] at (-2.1,-2.5) {$B$};

@@

  \node at (3,-3.5) {$c_2$};
  \node at (3,-4.5) {$c_1$};

  \node at (1.5,0.5) {$c_1$};
  \node at (1.5,-0.5) {$c_2$};

@@

\end{tikzpicture}
\end{center}
The top path is the $\Pi$ program
$(c_1~\oplus~c_2)~\odot~\swapp$ which deforms the
type $A$ by $c_1$, deforms the type $B$ by $c_2$, and deforms the
resulting space by a twist that exchanges the two injections into the
sum type. The bottom path performs the twist first and then deforms
the type $A$ by $c_1$ and the type $B$ by $c_2$ as before. If one
could imagine the paths are physical wires and the deformations $c_1$
and $c_2$ as arbitrary deformations on these wires then, holding the
points $A$, $B$, $C$, and $D$ fixed, it is possible to rotate the top
part of the diagram to become identical to the bottom one. In other
words, there exists a higher-level deformation of the program
$(c_1~\oplus~c_2)~\odot~\swapp$ to the program
$\swapp \odot (c_2~\oplus~c_1)$.
